19Q2

### 2019-07-08 19Q2.Rmd
phased out papaja for manuscript writing

DepMap 19Q2 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r depmap_metadata}
ccl_meta <- read.delim("./data_munging/19Q2/sample_info.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_meta <- select(ccl_meta, DepMap_ID, stripped_cell_line_name, CCLE_name, disease)
colnames(ccl_meta) <- c("DepMap_ID", "Cell_line", "CCLE_name", "disease")
# Remove two rows that contain cell lines with two different histologies
# The histologies that are kept are verified through matching in Cellosaurus and Cell Model Passports
# KMH2, haematopoietic_and_lymphoid: https://web.expasy.org/cellosaurus/CVCL_1330, https://cellmodelpassports.sanger.ac.uk/passports/SIDM01018
# MS1, lung: https://web.expasy.org/cellosaurus/CVCL_1429, https://cellmodelpassports.sanger.ac.uk/passports/SIDM00245
ccl_meta <- subset(ccl_meta, CCLE_name != "KMH2_THYROID" & CCLE_name != "MS1_SKIN")
ccl_meta <- subset(ccl_meta, !grepl("MERGED", CCLE_name, fixed = TRUE))
ccl_diseases <- as.character(unique(ccl_meta$disease))[as.character(unique(ccl_meta$disease)) != " "]
CCLE_histology <- tolower(gsub("^[a-z0-9]*_", "", ccl_meta$CCLE_name, ignore.case = TRUE))
ccl_meta <- add_column(ccl_meta, CCLE_histology = CCLE_histology, .after = "CCLE_name")
ccl_meta$disease <- with(ccl_meta, ifelse(disease == " " & CCLE_histology %in% ccl_diseases, as.character(CCLE_histology), as.character(disease)))
ccl_meta$disease <- with(ccl_meta, ifelse(disease == "ovary", "ovarian",
                                   ifelse(disease == " " & CCLE_histology == "vulva", "vulva",
                                   ifelse(disease == " " & CCLE_histology == "stomach", "gastric",
                                   ifelse(disease == " " & CCLE_histology == "upper_aerodigestive_tract", "upper_aerodigestive_tract",
                                   ifelse(disease == " " & CCLE_histology == "large_intestine", "colorectal",
                                   ifelse(disease == " " & CCLE_histology == "biliary_tract", "bile_duct",
                                   ifelse(disease == " " & CCLE_histology == "oesophagus", "esophagus",
                                   ifelse(disease == " " & CCLE_histology == "testis", "embryo", disease)))))))))
```

@@

```{r fusion_calls, eval=TRUE}
fusions <- read.delim("./data_munging/19Q2/CCLE_fusions.csv.gz", header = TRUE, sep = ",")
fusions_bcr_abl1 <- filter(fusions, X.FusionName %in% c("BCR--ABL1", "ABL1--BCR"))
fusions$LeftGene <- gsub(" .*", "", fusions$LeftGene)
fusions$RightGene <- gsub(" .*", "", fusions$RightGene)
fusions$X.FusionName <- gsub("--", "-", fusions$X.FusionName)
fusions <- select(fusions, DepMap_ID, X.FusionName, LeftGene, RightGene)
fusions <- as.data.frame(lapply(fusions, function(x) gsub("IGH@|IGH-@", "IGH", x)))
colnames(fusions) <- c("DepMap_ID", "FusionPair", "LeftGene", "RightGene")
fusions <- merge(fusions, ccl_meta, by = "DepMap_ID", all.x = TRUE)
```

@@

ccle <- MergeOmicsG2P(database = "CCLE", dataframe = ccle_all, maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./data_munging/19Q2/ccle_data_19Q2.rds")
write.csv.gz(ccle %>% drop_na(Score), "./data_munging/19Q2/ccle_data_19Q2.csv.gz")

ctrp <- MergeOmicsG2P(database = "CTRP", dataframe = ctrp_all, maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./../htscreens_giant_files/ctrp_data_19Q2.rds")
write.csv.gz(ctrp %>% drop_na(Score), "./../htscreens_giant_files/ctrp_data_19Q2.csv.gz")

gdsc <- MergeOmicsG2P(database = "GDSC", dataframe = gdsc_all, maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./data_munging/19Q2/gdsc_data_19Q2.rds")
write.csv.gz(gdsc %>% drop_na(Score), "./../htscreens_giant_files/gdsc_data_19Q2.csv.gz")

@@

```{r, eval=FALSE}
crispr_all <- read.delim("./data_munging/19Q2/Achilles_gene_effect_20190514.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr_all) <- gsub(" .*", "", colnames(crispr_all))
colnames(crispr_all)[1] <- "DepMap_ID"
maf_changes_df <- readRDS("./data_munging/19Q2/maf_changes_df_19Q2.rds")
maf_changes_filt <- filter(maf_changes_df, Gene_Protein_Change %in% unique(as.character(g2p_prot$Gene_Protein_Change)))

# CCLE
crispr_prot_grid <- expand.grid("DepMap_ID" = unique(crispr_all$DepMap_ID), "Gene_Protein_Change" = unique(maf_changes_filt$Gene_Protein_Change))
crispr_prot <- merge(crispr_all, crispr_prot_grid, by = "DepMap_ID", all = TRUE)
crispr_prot <- merge(crispr_prot, maf_changes_filt, by = c("DepMap_ID", "Gene_Protein_Change"), all = TRUE)
crispr_prot$Change_Status <- ifelse(!is.na(crispr_prot$Protein_Change), "Yes", "No")
crispr_prot$G2P_Protein_Change <- ifelse(crispr_prot$Gene_Protein_Change %in% unique(as.character(g2p_prot$Gene_Protein_Change)), "Yes", "No")
crispr_prot <- merge(crispr_prot, unique(select(g2p_prot, Gene_Protein_Change, Direction_Curated)), by = "Gene_Protein_Change", all.x = TRUE)
crispr_prot$Database <- "CRISPR"

@@

saveRDS(crispr_prot, "./data_munging/rds/protein_changes_merged.rds")
```

```{r}
crispr_prot <- readRDS("./data_munging/rds/protein_changes_merged.rds")
```

@@

# WATERFALL/GSEA: PanCAn

Consider looking at GSEA rankings.

>>>>

DepMap 19Q2 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r depmap_metadata, eval=FALSE}
ccl_meta <- read.delim("./data_munging/19Q2/sample_info.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_meta <- select(ccl_meta, DepMap_ID, stripped_cell_line_name, CCLE_name, disease)
colnames(ccl_meta) <- c("DepMap_ID", "Cell_line", "CCLE_name", "disease")
# Remove two rows that contain cell lines with two different histologies
# The histologies that are kept are verified through matching in Cellosaurus and Cell Model Passports
# KMH2, haematopoietic_and_lymphoid: https://web.expasy.org/cellosaurus/CVCL_1330, https://cellmodelpassports.sanger.ac.uk/passports/SIDM01018
# MS1, lung: https://web.expasy.org/cellosaurus/CVCL_1429, https://cellmodelpassports.sanger.ac.uk/passports/SIDM00245
ccl_meta <- subset(ccl_meta, CCLE_name != "KMH2_THYROID" & CCLE_name != "MS1_SKIN")
ccl_meta <- subset(ccl_meta, !grepl("MERGED", CCLE_name, fixed = TRUE))
ccl_diseases <- as.character(unique(ccl_meta$disease))[as.character(unique(ccl_meta$disease)) != " "]
CCLE_histology <- tolower(gsub("^[a-z0-9]*_", "", ccl_meta$CCLE_name, ignore.case = TRUE))
ccl_meta <- add_column(ccl_meta, CCLE_histology = CCLE_histology, .after = "CCLE_name")
ccl_meta$disease <- with(ccl_meta, ifelse(disease == " " & CCLE_histology %in% ccl_diseases, as.character(CCLE_histology), as.character(disease)))
ccl_meta$disease <- with(ccl_meta, ifelse(disease == "ovary", "ovarian",
                                   ifelse(disease == " " & CCLE_histology == "vulva", "vulva",
                                   ifelse(disease == " " & CCLE_histology == "stomach", "gastric",
                                   ifelse(disease == " " & CCLE_histology == "upper_aerodigestive_tract", "upper_aerodigestive_tract",
                                   ifelse(disease == " " & CCLE_histology == "large_intestine", "colorectal",
                                   ifelse(disease == " " & CCLE_histology == "biliary_tract", "bile_duct",
                                   ifelse(disease == " " & CCLE_histology == "oesophagus", "esophagus",
                                   ifelse(disease == " " & CCLE_histology == "testis", "embryo", disease)))))))))

write.csv(ccl_meta, "./data_munging/19Q2/sample_info_edit_20190708.csv")
```

```{r}
ccl_meta <- read.delim("./data_munging/19Q2/sample_info_edit_20190708.csv", sep = ",", header = TRUE, row.names = 1)
```

@@

```{r fusion_calls, eval=FALSE}
fusions <- read.delim("./data_munging/19Q2/CCLE_fusions.csv.gz", header = TRUE, sep = ",")
fusions_bcr_abl1 <- filter(fusions, X.FusionName %in% c("BCR--ABL1", "ABL1--BCR"))
fusions$LeftGene <- gsub(" .*", "", fusions$LeftGene)
fusions$RightGene <- gsub(" .*", "", fusions$RightGene)
fusions$X.FusionName <- gsub("--", "-", fusions$X.FusionName)
fusions <- select(fusions, DepMap_ID, X.FusionName, LeftGene, RightGene)
fusions <- as.data.frame(lapply(fusions, function(x) gsub("IGH@|IGH-@", "IGH", x)))
colnames(fusions) <- c("DepMap_ID", "FusionPair", "LeftGene", "RightGene")
fusions <- merge(fusions, ccl_meta, by = "DepMap_ID", all.x = TRUE)
write.csv.gz(fusions, "./data_munging/19Q2/CCLE_fusions_edit_20190708.csv.gz")
```

```{r}
fusions <- read.delim("./data_munging/19Q2/CCLE_fusions_edit_20190708.csv.gz", header = TRUE, sep = ",")
```

@@

ccle <- MergeOmicsG2P(database = "CCLE", dataframe = ccle_all, maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./data_munging/19Q2/ccle_data_19Q2.rds")
# write.csv.gz(ccle %>% drop_na(Score), "./data_munging/19Q2/ccle_data_19Q2.csv.gz")

ctrp <- MergeOmicsG2P(database = "CTRP", dataframe = ctrp_all, maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./../htscreens_giant_files/ctrp_data_19Q2.rds")
# write.csv.gz(ctrp %>% drop_na(Score), "./../htscreens_giant_files/ctrp_data_19Q2.csv.gz")

gdsc <- MergeOmicsG2P(database = "GDSC", dataframe = gdsc_all, maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./data_munging/19Q2/gdsc_data_19Q2.rds")
# write.csv.gz(gdsc %>% drop_na(Score), "./../htscreens_giant_files/gdsc_data_19Q2.csv.gz")

@@

```{r merge_crispr_fus, eval=FALSE}
fusions_filt <- filter(fusions, FusionPair %in% unique(as.character(g2p_fus$FusionPair)))

crispr_fus_grid <- expand.grid("DepMap_ID" = unique(crispr$DepMap_ID), "FusionPair" = unique(fusions_filt$FusionPair))
crispr_fus <- merge(merge(crispr, crispr_fus_grid, by = "DepMap_ID", all = TRUE), fusions_filt, by = c("DepMap_ID", "CCLE_name", "CCLE_histology", "Cell_line", "FusionPair"), all = TRUE)
crispr_fus$Fusion_Status <- ifelse(!is.na(crispr_fus$LeftGene), "Yes", "No")
crispr_fus$Drug_FusionPair <- ifelse(is.na(crispr_fus$Drug) | is.na(crispr_fus$FusionPair), NA, paste0(crispr_fus$Drug, "_", crispr_fus$FusionPair))
crispr_fus$G2P_Drug_FusionPair <- ifelse(crispr_fus$Drug_FusionPair %in% unique(as.character(g2p_fus$Drug_FusionPair)), "Yes", "No")
crispr_fus$G2P_Fusion <- ifelse(crispr_fus$FusionPair %in% unique(as.character(g2p_fus$FusionPair)), "Yes", "No")
crispr_fus$Database <- "CRISPR"

colnames(crispr_fus)[colnames(crispr_fus) == "AUC_IC50"] <- "Score"
saveRDS(crispr_fus, "./data_munging/rds/crispr_fusions.rds")
```

```{r merge_crispr_prot, eval=FALSE}
maf_changes_df <- readRDS("./data_munging/19Q2/maf_changes_df_19Q2.rds")
maf_changes_filt <- filter(maf_changes_df, Gene_Protein_Change %in% unique(as.character(g2p_prot$Gene_Protein_Change)))

# CCLE
crispr_prot_grid <- expand.grid("DepMap_ID" = unique(crispr$DepMap_ID), "Gene_Protein_Change" = unique(maf_changes_filt$Gene_Protein_Change))
crispr_prot <- merge(crispr, crispr_prot_grid, by = "DepMap_ID", all = TRUE)
crispr_prot <- merge(crispr_prot, maf_changes_filt, by = c("DepMap_ID", "Gene_Protein_Change"), all = TRUE)
crispr_prot$Change_Status <- ifelse(!is.na(crispr_prot$Protein_Change), "Yes", "No")
crispr_prot$G2P_Protein_Change <- ifelse(crispr_prot$Gene_Protein_Change %in% unique(as.character(g2p_prot$Gene_Protein_Change)), "Yes", "No")
crispr_prot <- merge(crispr_prot, unique(select(g2p_prot, Gene_Protein_Change, Direction_Curated)), by = "Gene_Protein_Change", all.x = TRUE)
crispr_prot$Database <- "CRISPR"

@@

saveRDS(crispr_prot, "./data_munging/rds/crispr_protein_changes.rds")
```

```{r}
crispr_fus <- readRDS("./data_munging/rds/crispr_fusions.rds")
crispr_prot <- readRDS("./data_munging/rds/crispr_protein_changes.rds")
```

@@
